High-order Solution Transfer between Curved Triangular Meshes

### 2020-09-29 doc/paper.tex
Fixing typo: "straight side" -> "straight sided".
Major pass on the introduction.

Solution transfer between meshes is a common problem that
has practical applications to many methods in
computational physics. For example, by allowing the underyling
computational domain to change during a simulation, computational
effort can be focused dynamically to resolve sensitive features
of a numerical solution. Mesh adaptivity, this in-flight change in the mesh,
requires translating the numerical solution from the old mesh to the new,
i.e. solution transfer. As another example, Lagrangian or particle-based
methods treat each node in the mesh as a particle and so with each timestep the
mesh travels \emph{with} the fluid.
However, over (typically limited) time the mesh
becomes distorted and suffers a loss in element quality which causes
catastrophic loss in the accuracy of computation. To overcome this, the
domain must be remeshed or rezoned and the solution must be
transferred (remapped) onto the new mesh configuration. A few scenarios
where solution transfer is necessary will be considered below
to motivate the ``black box'' solution transfer algorithm.

Solution transfer is needed when a solution
(approximated by a discrete field) is known on a \emph{donor} mesh and must
be transferred to a \emph{target} mesh. In many applications, the field
must be conserved for physical reasons, e.g. mass or energy cannot leave or
enter the system, hence the focus on \emph{conservative} solution transfer
(typically using Galerkin or \(L_2\)-minimizing methods).

When pointwise interpolation is used to transfer a solution, quantities with
physical meaning (e.g. mass, concentration, energy) may not be conserved.
To address this, there have been many explorations of
\emph{conservative interpolation}. Since solution transfer is so commonly
needed in physical applications, this problem of conservative interpolation
has been considered already for straight sided meshes.
However, both to allow for greater
geometric flexibility and for high order of convergence, this paper will
consider the case of curved isoparametric\footnote{I.e. the degree of the
discrete field on the mesh is same as the degree of the shape functions that
determine the mesh.} meshes. High-order methods
have the ability to produce highly accurate solutions
with low dissipation and low dispersion error, however, they typically
require curved meshes. In addition, curved elements have
advantages in problems that involve geometries that change over time,
such as flapping flight or fluid-structure interactions.

The \emph{common refinement} approach in
\cite{Jiao2004} is used to compare several methods for solution transfer across
two meshes. However, the problem of constructing a common refinement is
not discussed there. The problem of constructing such a refinement is
considered in \cite{Farrell2009, Farrell2011} (called a supermesh by
the authors).
However, the solution transfer becomes considerably more challenging for curved
meshes. For a sense of the difference between the straight sided and curved
cases, consider the problem of intersecting an element from the donor mesh
with an element from the target mesh. If the elements are triangles, the
intersection is either a convex polygon or has measure zero. If the elements
are curved, the intersection can be non-convex and can even split into
multiple disjoint regions. In \cite{Qiu2018}, a related problem is considered
where squares are allowed to curve along characteristics and the solution is
remapped onto a regular square grid.

Lagrangian methods treat each point in the
physical domain as a ``particle'' which moves along a characteristic curve
over time and then monitor values associated with the particle (heat / energy,
velocity, pressure, density, concentration, etc.). They are an effective way
to solve PDEs, even with higher order or non-linear terms \cite{MR3023731}.
This approach transforms the numerical solution of PDEs into a family of
numerical solutions to many independent ODEs. It allows the use of familiar
and well understood ODE solvers. In addition, Lagrangian methods often
have less restrictive conditions on time steps than Eulerian
methods\footnote{In Eulerian methods, the mesh is fixed.}.
When solving PDEs on unstructured meshes
with Lagrangian methods, the nodes move (since they
are treated like particles) and the mesh ``travels''.
A flow-based change to a mesh can cause problems if it
distorts the mesh until the element
quality is too low in some mesh elements. Over enough time, the mesh can
even tangle (i.e. elements begin to overlap).

To deal with distortion, one can allow the mesh to adapt in between time
steps. In addition to improving mesh quality, mesh
adaptivity can be used to dynamically focus computational effort to resolve
sensitive features of a numerical solution. From \cite{Iske2004}
\begin{quote}
{\small In order to balance the method's approximation quality and its
computational costs effectively, adaptivity is an essential requirement,
especially when modelling multiscale phenomena.}
\end{quote}
For more on mesh adaptivity, see \cite{Babuska1978, Peraire1987, Pain2001}.

In either case, the change in the
mesh between time steps requires transferring a known solution on the
discarded mesh to the mesh produced by the remeshing process. Without
the ability to change the mesh, Lagrangian methods (or, more generally,
ALE\footnote{ALE, or Arbitrary Lagrangian-Eulerian methods are not constrained
by a fixed computational mesh (the Eulerian approach) or a fixed fluid flow
(the Lagrangian approach) and typically combine the benefits of both approaches
to create composite methods} \cite{Hirt1974}) would not be useful, since after
a limited time the mesh will distort.

To allow for greater geometric flexibility and for high order of convergence,
curved mesh elements can be used in the finite element method. Though the
complexity of a method can steeply rise when allowing curved elements, the
trade for high-order convergence can be worth it. (See \cite{Wang2013} for
more on high-order CFD methods.) Curved meshes can typically
represent a given geometry with far fewer elements than a straight sided mesh.
The increase in accuracy also allows for the use of fewer elements, which
in turn can also facilitate a reduction in the overall computation time.

\begin{figure}
  \includegraphics{element_distortion.pdf}
  \centering
  \captionsetup{width=.75\linewidth}
  \caption{Movement of nodes in a quadratic element under distortion caused
    by particle motion along the velocity field \(\left[ y^2 \; 1 \right]^T\)
    from \(t = 0\) to \(t = 1\) with \(\Delta t = 1/2\). The green curves
    represent the characteristics that each node travels along.}
  \label{fig:element-distortion}
\end{figure}

Even if the domain has no inherent curvature, high-order (degree \(p\)) shape
functions allow for order \(p + 1\) convergence, which is desirable in
its own right. However, even in such cases, a Lagrangian method must either
curve the mesh or information about the flow of the geometry will be lost.
Figure~\ref{fig:element-distortion} shows what happens to a given
quadratic element as the nodes move along the
characteristics given by the velocity field
\(\left[ \begin{array}{c c} y^2 & 1 \end{array}\right]^T\).
This element uses the triangle vertices and edge
midpoints to determine the shape functions. However, as the nodes move
with the flow, the midpoints are no longer on the lines connecting
the vertex nodes. To allow the mesh to more accurately represent the
solution, the edges can instead curve so that the midpoint nodes remain
halfway between (i.e. half of the parameter space) the vertex nodes along
an edge.

In multiphysics simulations, a problem is partitioned into physical components.
This partitioning can apply to both the physical domain (e.g. separating a
solid and fluid at an interface) and the simulation solution itself (e.g.
solving for pressure on one mesh and velocity on another). Each (multi)physics
component is solved for on its own mesh. When the components interact, the
simulation solution must be transferred between those meshes.

In a similar category of application, solution transfer enables the comparison
of solutions defined on different meshes. For example, if a reference
solution is known on a very fine special-purpose mesh, the error can be
computed for a coarse mesh by transferring the solution from the fine
mesh and taking the difference. Or, if the same method is used on
different meshes of the same domain, the resulting computed solutions can
be compared via solution transfer. Or, if two different methods use two
different meshes of the same domain, a solution may need to be transferred
across these meshes for comparison.

@@

\begin{figure}
  \includegraphics{main_figure29.pdf}
  \centering
  \captionsetup{width=.75\linewidth}
  \caption{Partially overlapping meshes on a near identical domain. Both are
    linear meshes that approximate the unit disc in \(\reals^2\). The outermost
    columns show how the domain of each mesh can be expanded so they agree.}
  \label{fig:partially-overlapping}
\end{figure}

@@

The case of partially overlapping meshes can be addressed in particular
cases (i.e. with more information). For example, consider a problem
defined on \(\Omega = \reals^2\) and solution
that tends towards zero as points tend to infinity. A typical approach
may be to compute the solution on a circle of large enough radius and
consider the numerical solution to be zero outside the circle.
Figure~\ref{fig:partially-overlapping} shows how solution transfer could be
performed in such cases when the meshes partially overlap: construct a
simple region containing both computational domains and then mesh the
newly introduced area. However, the assumption that the numerical solution
is zero in the newly introduced area is very specific and
a similar approach may not apply in other cases of partial overlap.

Some attempts (\cite{Berger1987, Chesshire1994, Cai1999}) have been
made to interpolate fluxes between overlapping meshes. These perform
an interpolation on the region common to both meshes and then numerically
solve the PDE to determine the values on the uncovered elements.

When solving some PDEs with viscoelastic fields, additional nodes must be
tracked at quadrature points. As the mesh travels, these quadrature points
must be moved as well in a way that is consistent with the standard
nodes that define the element. The solution transfer method described here
does not provide a way either to determine quadrature points on a
target mesh or to map the field onto known quadrature points.

@@

Convergence results are shown in Figure~\ref{fig:composite-errors}
and confirm the expected orders. Since \(\zeta_1\) is a cubic polynomial,
one might expect the solution transfer on the cubic mesh to be \emph{exact}.
This expectation would hold in the superparametric case (i.e. where the
elements are straight sided but the shape functions are higher degree).
However, when the elements aren't straight side the isoparametric mapping
changes the underlying function space.

>>>>

Solution transfer between meshes is a common problem that
has practical applications to many methods in
computational physics. For example, by allowing the underyling
computational domain to change during a simulation, computational
effort can be focused dynamically to resolve relevant features
of a numerical solution. This so-called mesh adaptivity typically
requires translating the numerical solution from the old mesh to the new,
i.e. solution transfer \cite{Babuska1978,Peraire1987,Pain2001,Iske2004}. As another example, Lagrangian or particle-based
methods treat each node in the mesh as a particle and so with each timestep the
mesh travels \emph{with} the fluid \cite{MR3023731}.
However, over (typically short) time the mesh
becomes distorted and suffers a loss in element quality which eventually causes
inaccurate solutions. To overcome this, the
domain must be remeshed or rezoned and the solution must be
transferred (remapped) onto the new mesh configuration. A related more general
class of methods is the Arbitrary Lagrangian-Eulerian (ALE) method, which
are not constrained by a fixed computational mesh (the Eulerian approach) or
by a fixed fluid flow (the Lagrangian approach) and typically combine the benefits
of both approaches \cite{Hirt1974,MR3302403}.

Solution transfer is needed when a solution
(approximated by a discrete field) is known on a \emph{donor} mesh and must
be transferred to a \emph{target} mesh.
When pointwise interpolation is used to transfer a solution, quantities with
physical meaning (e.g. mass, concentration, energy) may not be conserved.
In many applications, the field
must be conserved for physical reasons, e.g. mass or energy cannot leave or
enter the system, hence this paper focuses on \emph{conservative} solution transfer
(typically using Galerkin or \(L_2\)-minimizing methods).

%% When pointwise interpolation is used to transfer a solution, quantities with
%% physical meaning (e.g. mass, concentration, energy) may not be conserved.
%% To address this, there have been many explorations of
%% \emph{conservative interpolation}. Since solution transfer is so commonly
%% needed in physical applications, this ...

The problem of conservative interpolation
has been considered previously for straight sided meshes.
However, both to allow for greater
geometric flexibility and for high order of convergence, this paper
addresses the case of curved isoparametric meshes. This is based on the
recent interest in so-called high-order methods \cite{Wang2013}, which have the ability to
produce highly accurate solutions with low dissipation and low dispersion error.
However, these methods typically require curved meshes to obtain the high accuracy.

%% In addition, curved elements have advantages in problems that involve geometries that change over time,
%% such as flapping flight or fluid-structure interactions.

The \emph{common refinement} approach in
\cite{Jiao2004} is used to compare several methods for solution transfer across
two meshes. The problem of constructing such a refinement is
considered in \cite{Farrell2009, Farrell2011} (called a supermesh by
the authors).
However, the solution transfer becomes considerably more challenging for curved
meshes. For a sense of the difference between the straight sided and curved
cases, consider the problem of intersecting an element from the donor mesh
with an element from the target mesh. If the elements are triangles, the
intersection is either a convex polygon or has measure zero. If the elements
are curved, the intersection can be non-convex and can even split into
multiple disjoint regions. In \cite{Qiu2018}, a related problem is considered
where squares are allowed to curve along characteristics and the solution is
remapped onto a regular square grid.

%% Lagrangian methods treat each point in the
%% physical domain as a ``particle'' which moves along a characteristic curve
%% over time and then monitor values associated with the particle (heat / energy,
%% velocity, pressure, density, concentration, etc.). They are an effective way
%% to solve PDEs, even with higher order or non-linear terms \cite{MR3023731}.
%% This approach transforms the numerical solution of PDEs into a family of
%% numerical solutions to many independent ODEs. It allows the use of familiar
%% and well understood ODE solvers. In addition, Lagrangian methods often
%% have less restrictive conditions on time steps than Eulerian
%% methods\footnote{In Eulerian methods, the mesh is fixed.}.
%% When solving PDEs on unstructured meshes
%% with Lagrangian methods, the nodes move (since they
%% are treated like particles) and the mesh ``travels''.
%% A flow-based change to a mesh can cause problems if it
%% distorts the mesh until the element
%% quality is too low in some mesh elements. Over enough time, the mesh can
%% even tangle (i.e. elements begin to overlap).

%% To deal with distortion, one can allow the mesh to adapt in between time
%% steps. In addition to improving mesh quality, mesh
%% adaptivity can be used to dynamically focus computational effort to resolve
%% sensitive features of a numerical solution. From \cite{Iske2004}
%% \begin{quote}
%% {\small In order to balance the method's approximation quality and its
%% computational costs effectively, adaptivity is an essential requirement,
%% especially when modelling multiscale phenomena.}
%% \end{quote}
%% For more on mesh adaptivity, see \cite{Babuska1978, Peraire1987, Pain2001}.

%% In either case, the change in the
%% mesh between time steps requires transferring a known solution on the
%% discarded mesh to the mesh produced by the remeshing process. Without
%% the ability to change the mesh, Lagrangian methods (or, more generally,
%% ALE\footnote{ALE, or Arbitrary Lagrangian-Eulerian methods are not constrained
%% by a fixed computational mesh (the Eulerian approach) or a fixed fluid flow
%% (the Lagrangian approach) and typically combine the benefits of both approaches
%% to create composite methods} \cite{Hirt1974}) would not be useful, since after
%% a limited time the mesh will distort.

%% To allow for greater geometric flexibility and for high order of convergence,
%% curved mesh elements can be used in the finite element method. Though the
%% complexity of a method can steeply rise when allowing curved elements, the
%% trade for high-order convergence can be worth it. (See \cite{Wang2013} for
%% more on high-order CFD methods.) Curved meshes can typically
%% represent a given geometry with far fewer elements than a straight sided mesh.
%% The increase in accuracy also allows for the use of fewer elements, which
%% in turn can also facilitate a reduction in the overall computation time.

%% \begin{figure}
%%   \includegraphics{element_distortion.pdf}
%%   \centering
%%   \captionsetup{width=.75\linewidth}
%%   \caption{Movement of nodes in a quadratic element under distortion caused
%%     by particle motion along the velocity field \(\left[ y^2 \; 1 \right]^T\)
%%     from \(t = 0\) to \(t = 1\) with \(\Delta t = 1/2\). The green curves
%%     represent the characteristics that each node travels along.}
%%   \label{fig:element-distortion}
%% \end{figure}

%% Even if the domain has no inherent curvature, high-order (degree \(p\)) shape
%% functions allow for order \(p + 1\) convergence, which is desirable in
%% its own right. However, even in such cases, a Lagrangian method must either
%% curve the mesh or information about the flow of the geometry will be lost.
%% Figure~\ref{fig:element-distortion} shows what happens to a given
%% quadratic element as the nodes move along the
%% characteristics given by the velocity field
%% \(\left[ \begin{array}{c c} y^2 & 1 \end{array}\right]^T\).
%% This element uses the triangle vertices and edge
%% midpoints to determine the shape functions. However, as the nodes move
%% with the flow, the midpoints are no longer on the lines connecting
%% the vertex nodes. To allow the mesh to more accurately represent the
%% solution, the edges can instead curve so that the midpoint nodes remain
%% halfway between (i.e. half of the parameter space) the vertex nodes along
%% an edge.

%% In multiphysics simulations, a problem is partitioned into physical components.
%% This partitioning can apply to both the physical domain (e.g. separating a
%% solid and fluid at an interface) and the simulation solution itself (e.g.
%% solving for pressure on one mesh and velocity on another). Each (multi)physics
%% component is solved for on its own mesh. When the components interact, the
%% simulation solution must be transferred between those meshes.

%% In a similar category of application, solution transfer enables the comparison
%% of solutions defined on different meshes. For example, if a reference
%% solution is known on a very fine special-purpose mesh, the error can be
%% computed for a coarse mesh by transferring the solution from the fine
%% mesh and taking the difference. Or, if the same method is used on
%% different meshes of the same domain, the resulting computed solutions can
%% be compared via solution transfer. Or, if two different methods use two
%% different meshes of the same domain, a solution may need to be transferred
%% across these meshes for comparison.

@@

In addition to conservation, there are many other considerations for a
solution transfer method that are important in various applications. For
example, since our focus is on high-order methods, these typically produce
oscillations in the solutions that can be non-physical and break the numerical
solvers. This can happen for the standard Galerkin projections that we
consider here, which means that e.g. discontinuities such as shocks in the
solution could become nonphysical after our solution transfer. Various
solutions to this have been proposed before, often referred to as limiting or
bound/monotonicity preserving methods, see e.g. \cite{MR3302403,Anderson2018}
where several approaches are studied including a high-order extension of the
flux-corrected transport (FCT) method. Other considerations for a solution
transfer scheme include the preservation of additional constraints, such
as the divergence or the curl of the solution.

@@

%%%%%%%%%%%%%%%
%% Overlapping meshes - exclude for now

%% \begin{figure}
%%   \includegraphics{main_figure29.pdf}
%%   \centering
%%   \captionsetup{width=.75\linewidth}
%%   \caption{Partially overlapping meshes on a near identical domain. Both are
%%     linear meshes that approximate the unit disc in \(\reals^2\). The outermost
%%     columns show how the domain of each mesh can be expanded so they agree.}
%%   \label{fig:partially-overlapping}
%% \end{figure}

%% The case of partially overlapping meshes can be addressed in particular
%% cases (i.e. with more information). For example, consider a problem
%% defined on \(\Omega = \reals^2\) and solution
%% that tends towards zero as points tend to infinity. A typical approach
%% may be to compute the solution on a circle of large enough radius and
%% consider the numerical solution to be zero outside the circle.
%% Figure~\ref{fig:partially-overlapping} shows how solution transfer could be
%% performed in such cases when the meshes partially overlap: construct a
%% simple region containing both computational domains and then mesh the
%% newly introduced area. However, the assumption that the numerical solution
%% is zero in the newly introduced area is very specific and
%% a similar approach may not apply in other cases of partial overlap.

%% Some attempts (\cite{Berger1987, Chesshire1994, Cai1999}) have been
%% made to interpolate fluxes between overlapping meshes. These perform
%% an interpolation on the region common to both meshes and then numerically
%% solve the PDE to determine the values on the uncovered elements.

%%%%%%%%%%%%%%%

%% When solving some PDEs with viscoelastic fields, additional nodes must be
%% tracked at quadrature points. As the mesh travels, these quadrature points
%% must be moved as well in a way that is consistent with the standard
%% nodes that define the element. The solution transfer method described here
%% does not provide a way either to determine quadrature points on a
%% target mesh or to map the field onto known quadrature points.

@@

Convergence results are shown in Figure~\ref{fig:composite-errors}
and confirm the expected orders. Since \(\zeta_1\) is a cubic polynomial,
one might expect the solution transfer on the cubic mesh to be \emph{exact}.
This expectation would hold in the superparametric case (i.e. where the
elements are straight sided but the shape functions are higher degree).
However, when the elements aren't straight sided the isoparametric mapping
changes the underlying function space.
